---
title: "sc2marker with MCA-spleen (Mouse)"
author: "sc2marker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(sc2marker))
```

# load data

```{r SeuratObject}
mca.spleen
Idents(mca.spleen) <- "anno"
ids <- c("T cell(Spleen)", "Macrophage(Spleen)", "NK cell(Spleen)", "Marginal zone B cell(Spleen)")
mca.spleen@meta.data$annotation_simplified <- mca.spleen@meta.data$anno
```

# Simplify the cell annotation

```{r}
ids <- c("T cell(Spleen)", "Macrophage(Spleen)", "NK cell(Spleen)", "Marginal zone B cell(Spleen)")
mca.spleen@meta.data$annotation_simplified <- ifelse(mca.spleen@meta.data$annotation_simplified %in% ids,  mca.spleen@meta.data$annotation_simplified, "Other")
mca.spleen@meta.data$annotation_simplified[mca.spleen@meta.data$annotation_simplified == "T cell(Spleen)"] <- "T Cell"
mca.spleen@meta.data$annotation_simplified[mca.spleen@meta.data$annotation_simplified == "Marginal zone B cell(Spleen)"] <- "B Cell"
mca.spleen@meta.data$annotation_simplified[mca.spleen@meta.data$annotation_simplified == "NK cell(Spleen)"] <- "NK Cell"
mca.spleen@meta.data$annotation_simplified[mca.spleen@meta.data$annotation_simplified == "Macrophage(Spleen)"] <- "Macro"
Idents(mca.spleen) <- "annotation_simplified"
```

# Markers identification for NK cell

To run sc2marker you need to execute the following command, providing a clustered single cell data sets (as Seurat object), the cell type of interest and the antibody databased (IHC, ICC or Flow).

```{r NK cell markers}
nk.markers <- Detect_single_marker(mca.spleen, id = "NK Cell", category = "Flow", org = "mouse", min.pct = 0.1)
get_antibody(nk.markers, org = "mouse")
```

```{r message=F, warning=F, fig.height=9, fig.width=9}
# Ridge plot of top 9 markers of NK cells
plot_ridge(mca.spleen, id = "NK Cell", genes = nk.markers[1:9,]$gene, ncol = 3, assay = "RNA", aggr.other = F)
```


# Markers identification for B cell

```{r B cell markers}
b.markers <- Detect_single_marker(mca.spleen, id = "B Cell", category = "Flow", org = "mouse", min.pct = 0.1)
get_antibody(b.markers, org = "mouse")
```

```{r message=F, warning=F, fig.height=9, fig.width=9}
# Ridge plot of top 9 markers of B cells
plot_ridge(mca.spleen, id = "B Cell", genes = b.markers[1:9,]$gene, ncol = 3, assay = "RNA", aggr.other = F)
```



# Markers identification for T cell

```{r T cell markers}
t.markers <- Detect_single_marker(mca.spleen, id = "T Cell", category = "Flow", org = "mouse", min.pct = 0.1)
get_antibody(t.markers, org = "mouse")
```

```{r message=F, warning=F, fig.height=9, fig.width=9}
# Ridge plot of top 9 markers of B cells
plot_ridge(mca.spleen, id = "T Cell", genes = t.markers[1:9,]$gene, ncol = 3, assay = "RNA", aggr.other = F)
```


# Markers identification for Macrophages

```{r Macro cell markers}
Macro.markers <- Detect_single_marker(mca.spleen, id = "Macro", category = "Flow", org = "mouse", min.pct = 0.1)
get_antibody(Macro.markers, org = "mouse")
```

```{r message=F, warning=F, fig.height=9, fig.width=9}
# Ridge plot of top 9 markers of B cells
plot_ridge(mca.spleen, id = "Macro", genes = Macro.markers[1:9,]$gene, ncol = 3, assay = "RNA", aggr.other = F)
```

# Session Info

```{r}
sessionInfo()
```

